\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry}
\geometry{margin=1in}
\title{Recursive Drift Collapse Theorem}
\author{}
\date{}

\begin{document}
\maketitle

\section*{I. Theorem Statement}

Recursive information processing systems exhibit progressive degradation in output quality as recursion depth or density increases. There exists a critical threshold $\theta_c$ beyond which semantic coherence decreases, and the information gained per recursion step approaches zero.

Let:
\begin{itemize}
  \item $C(t)$ – Coherence function at recursion step $t$
  \item $\Delta I(t)$ – Information gain between recursive iterations
  \item $D_a(t)$ – Active recursion density
  \item $\theta_c$ – Critical recursion threshold
\end{itemize}

If:
\[
D_a(t) > \theta_c
\quad \Rightarrow \quad
\frac{d^2C}{dt^2} > 0 \quad \text{and} \quad \Delta I(t) \rightarrow 0
\]

This defines the onset of recursive drift collapse.

\section*{II. Collapse Mechanism}

\begin{enumerate}
  \item \textbf{Cumulative Drift:} Small output variations accumulate across iterations, reducing semantic integrity.
  \item \textbf{Reduced Information Gain:} $\Delta I(t)$ declines toward zero as outputs become less novel.
  \item \textbf{Accelerated Degradation:} Coherence loss accelerates near $\theta_c$, often exhibiting short-term oscillations before collapse.
\end{enumerate}

\section*{III. Collapse Conditions}

\begin{itemize}
  \item Recursion depth or density exceeds $\theta_c$
  \item No external semantic grounding is applied
  \item Input structure induces increased entropy over iterations
\end{itemize}

\section*{IV. Simulation Design}

\subsection*{Suggested Models}

\begin{itemize}
  \item Autoregressive transformers (e.g., GPT variants)
  \item Recursive neural networks (RNNs)
  \item Symbolic reasoning systems with feedback loops
\end{itemize}

\subsection*{Metrics}

\begin{itemize}
  \item Coherence scoring (e.g., BERTScore)
  \item Mutual Information estimation (MINE, CLUB)
  \item Cosine similarity of latent embeddings
  \item $\Delta I(t)$ trends over recursion steps
\end{itemize}

\subsection*{Visualization Techniques}

\begin{itemize}
  \item Embedding projections using UMAP or t-SNE
  \item Acceleration of coherence loss $\left(\frac{d^2C}{dt^2}\right)$
  \item Threshold indicators for $\Delta I(t) \rightarrow 0$
\end{itemize}

\section*{V. Human Cognition Analogy}

Human reasoning exhibits similar degradation under recursive load:
\begin{itemize}
  \item Breakdown in symbolic consistency during paradox tasks
  \item EEG entropy increases during recursive reasoning
  \item Working memory failures under semantic recursion pressure
\end{itemize}

\section*{VI. Applications}

\begin{itemize}
  \item AI failure detection in recursive text generation
  \item Modeling cognitive collapse under symbolic stress
  \item Recursion-safe design in language and symbolic systems
  \item Theoretical bounding of symbolic regressions
\end{itemize}

\section*{VII. Future Work}

\begin{itemize}
  \item Formal models of entropy buildup in recursion
  \item Design of stabilizing feedback mechanisms
  \item Exploration of drift-correction strategies in symbolic AI
\end{itemize}

\end{document}